---
title: "Assignment 4 - Recommendation System"
author: "Santiago Fraile"
date: "5/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Business Problem/Goal: The main goal of this project is to create a movie recommendation system that recommends movies to users; in this way we can have a win-win scenario, where the customer selects and watch the movie they would like to see and we will retain users on the platform to gain time for advertisement. 

# I have downloaded two csv datasets: movies.csv and ratings.csv. These datasets were downladed from Google Drive (https://drive.google.com/file/d/1Dn1BZD3YxgBQJSIjbfNnmCFlDW2jdQGD/view) and have saved them on my R work directory. ("C:/Users/HP/Desktop/R Projects/Module 4 - FDS/Recommendation System M4")

```{r}
getwd()

```

# We will make use of these four packages 'recommenderlab', 'ggplot2', 'data.table' and 'reshape2'

```{r}
library(recommenderlab)
library(ggplot2)
library(data.table)
library(reshape2)
```


# We will now retrieve the data from movies.csv into movie_data dataframe and rating.csv into rating_data. We will use the str() function to display information about the movie_data dataframe.

```{r}
setwd("C:/Users/HP/Desktop/R Projects")
movie_data <- read.csv("movies.csv", stringsAsFactors = FALSE)
rating_data <- read.csv("ratings.csv")
str(movie_data)
```

# We can also overview the summary of the movies and ratings, such as Mean, Media, Max, Min values.
```{r}
summary(movie_data)
summary(rating_data)
```

# Here you will see the first 6 rows of each dataset to understand the variables

```{r}
head(movie_data)
head(rating_data)
```

# Now I will describe the variables we have at movie_data

```{r}
text_tbl <- data.frame(
             ColumnName = c("Movie ID", "Title", "Genres"),
             BriefDecription = c( "ID assigned to movie", "Name of the movie", "If it is adventure, comdey, etc"),
             UnitofMeasurement = c( "Number", "Name", "Name")
)

print(text_tbl)
```

#and this are the once we have at rating_data

```{r}
text_tbl <- data.frame(
             ColumnName = c("User ID", "Movie ID", "Rating", "TimeStamp"),
             BriefDecription = c( "ID assigned to user", "ID assigned to movie", "Rating assigned to movie", "time of occurence "),
             UnitofMeasurement = c( "Number", "number", "Number", "Number")
)

print(text_tbl)
```

# In regards to data pre-processing we need to convert the genres present in the movie_data dataframe into a more usable format by the users.

```{r}
movie_genre <- as.data.frame(movie_data$genres, stringsAsFactors=FALSE)
library(data.table)
movie_genre2 <- as.data.frame(tstrsplit(movie_genre[,1], '[|]', 
                                   type.convert=TRUE), 
                         stringsAsFactors=FALSE) #DataFlair
colnames(movie_genre2) <- c(1:10)
list_genre <- c("Action", "Adventure", "Animation", "Children", 
                "Comedy", "Crime","Documentary", "Drama", "Fantasy",
                "Film-Noir", "Horror", "Musical", "Mystery","Romance",
                "Sci-Fi", "Thriller", "War", "Western")
genre_mat1 <- matrix(0,10330,18)
genre_mat1[1,] <- list_genre
colnames(genre_mat1) <- list_genre
for (index in 1:nrow(movie_genre2)) {
  for (col in 1:ncol(movie_genre2)) {
    gen_col = which(genre_mat1[1,] == movie_genre2[index,col]) #Author DataFlair
    genre_mat1[index+1,gen_col] <- 1
}
}
genre_mat2 <- as.data.frame(genre_mat1[-1,], stringsAsFactors=FALSE) #remove first row, which was the genre list
for (col in 1:ncol(genre_mat2)) {
  genre_mat2[,col] <- as.integer(genre_mat2[,col]) #convert from characters to integers
} 
str(genre_mat2)
```

#As a next step of Data Pre-processing we will create a ‘search matrix’ that will allow us to perform an easy search of the films by specifying the genre present in our list

```{r}
SearchMatrix <- cbind(movie_data[,1:2], genre_mat2[])
head(SearchMatrix)    #DataFlair
```

#Here are movies that have several genres, for example, Toy Story, which is an animated film also falls under the genres of Comedy, Fantasy, and Children. This applies to the majority of the films.For our movie recommendation system to make sense of our ratings through recommenderlabs, we have to convert our matrix into a sparse matrix one


```{r}
ratingMatrix <- dcast(rating_data, userId~movieId, value.var = "rating", na.rm=FALSE)
ratingMatrix <- as.matrix(ratingMatrix[,-1]) #remove userIds
#Convert rating matrix into a recommenderlab sparse matrix
ratingMatrix <- as(ratingMatrix, "realRatingMatrix")
ratingMatrix
```

#Let us now overview some of the important parameters that provide us various options for building recommendation systems for movies

```{r}
recommendation_model <- recommenderRegistry$get_entries(dataType = "realRatingMatrix")
names(recommendation_model)
```

```{r}
lapply(recommendation_model, "[[", "description")
```

#We will implement a single model in our R project – Item Based Collaborative Filtering

```{r}
recommendation_model$IBCF_realRatingMatrix$parameters
```

# Now we will do some graphs to visualize data
# Most Viewed Movies Visualization

```{r}
library(ggplot2)
movie_views <- colCounts(ratingMatrix) # count views for each movie
table_views <- data.frame(movie = names(movie_views),
                          views = movie_views) # create dataframe of views
table_views <- table_views[order(table_views$views,
                                 decreasing = TRUE), ] # sort by number of views
table_views$title <- NA
for (index in 1:10325){
  table_views[index,3] <- as.character(subset(movie_data,
                                         movie_data$movieId == table_views[index,1])$title)
}
table_views[1:6,]
```

#Now, we will visualize a bar plot for the total number of views of the top films. We will carry this out using ggplot2.

```{r}
ggplot(table_views[1:6, ], aes(x = title, y = views)) +
  geom_bar(stat="identity", fill = 'steelblue') +
  geom_text(aes(label=views), vjust=-0.3, size=3.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Total Views of the Top Films")

```

# Now lets do a heatmap for Movie Ratings. Will include first 25 rows and first 25 columns

```{r}
image(ratingMatrix[1:20, 1:25], axes = FALSE, main = "Heatmap of the first 25 rows and 25 columns")
```

# One of the key steps for executing a succesfull data preparation process was to select the useful data, and to do so we have done the following:We have set the threshold for the minimum number of users who have rated a film as 50. There was an error on the code provided in the reserach where the movie_rating was written as Movie_rating

```{r}
movie_ratings <- ratingMatrix[rowCounts(ratingMatrix) > 50,
                             colCounts(ratingMatrix) > 50]
movie_ratings
```
```{r}
minimum_movies<- quantile(rowCounts(movie_ratings), 0.98)
minimum_users <- quantile(colCounts(movie_ratings), 0.98)
image(movie_ratings[rowCounts(movie_ratings) > minimum_movies,
                     colCounts(movie_ratings) > minimum_users],
main = "Heatmap of the top users and movies")
```



